1. Importing libraries, functions and samples
```{r setup}
library(readxl)
library(tidyverse)
library(openxlsx)
library(adegenet)
library(pegas)

```

```{r}
# Function re-formats genind object into a convenient dataframe w/ diversity indices for each locus for each population
popSummLoc = function(genotypes, B = 100000){
  

  Population = NULL
  Locus = NULL
  N=NULL
  He = NULL
  Ho = NULL
  pHWE = NULL
  A = NULL
  
  bypop=seppop(genotypes, drop=TRUE)
  
  for(i in 1:length(bypop)){
    summ=summary(bypop[[i]], verbose=FALSE)
    NLoc = length(summ$loc.n.all)
    Population = c(Population, rep(levels(pop(bypop[[i]])),NLoc))
    Locus = c(Locus, locNames(bypop[[i]]))
    
    He = c(He, summ$Hexp)
    Ho = c(Ho, summ$Hobs)
    A = c(A, summ$loc.n.all)
    
    HWx = hw.test(bypop[[i]], B = B)  # using a fixed number of Monte Carlo replicates
    pHWE = rbind(pHWE, HWx)

  }
  return(cbind(data.frame(Population, Locus, He, Ho, A), as.data.frame(pHWE)))
  
}
```


```{r}
raw_data_all = read_xlsx("Data/Sample_data.xlsx", sheet = "Samples_data")
```




2. Create genind object with 20 microsatelite markers

```{r}
gendata_PCA <-  df2genind(raw_data_all[c("Cl147",	"Cl233",	"Cl264",	"Cl274",	"Cl285",	"Cl291",	"Cl308",	"Cl330",	"Cl344","Cl366",	"Cl375",	"Cl408",	"Cl507",	"Cl527",	"Cl380","Lup02",	"Lup13",	"Lup20",	"Lup21",	"Lup23")], sep = " ", NA.char = "-9 -9")
gendata_PCA@other <-  raw_data_all[,1:10]
indNames(gendata_PCA) <- raw_data_all$SampleID
pop(gendata_PCA) <-  raw_data_all$Classified_PCA
unique(pop(gendata_PCA))
```

3. Different explorations with Principal component analysis (PCA)

3.1 All samples

a) All samples
```{r}
x.pca <- dudi.pca(tab(gendata_PCA, NA.method = "mean"), scannf = FALSE, scale = FALSE)
s.class(x.pca$li, pop(gendata_PCA), col = hybridpal()(16))


```
b) Only dog and wolf samples (no hybrids)
```{r}
gendata_Country <- gendata_PCA[pop(gendata_PCA) %in% c ("Dog", "W-Slovenia", "W-Croatia", "W-BIH", "W-Serbia", "W-Kosovo", "W-Montenegro", "W-N.Macedonia", "W-Bulgaria", "W-Italy", "W-Germany", "W-Romania")]

x.pca <- dudi.pca(tab(gendata_Country, NA.method = "mean"), scannf = FALSE, scale = FALSE)
s.class(x.pca$li, pop(gendata_Country), col = hybridpal()(12))

```
We see that with PCA we cannot  distinguish well between different (sub)populations. Because of this, for our analyses PCA is not useful to look at all the samples together. We will try comparing samples from only few or two (sub)populations.

3.2. Comparison of different (sub)populations

#3.2.1) Different explorations with dogs, hybrids, Slovenian and Croatian samples

With hybrid samples
```{r}
gendata_Country <- gendata_PCA[pop(gendata_PCA) %in% c ("Dog", "F1","BC1w", "BC2w", "W-Slovenia", "W-Croatia")]

x.pca <- dudi.pca(tab(gendata_Country, NA.method = "mean"), scannf = FALSE, scale = FALSE)
s.class(x.pca$li, pop(gendata_Country), col = hybridpal()(11))

```
Without hybrid samples
```{r}
gendata_Country <- gendata_PCA[pop(gendata_PCA) %in% c ("Dog",  "W-Slovenia", "W-Croatia")] 

x.pca <- dudi.pca(tab(gendata_Country, NA.method = "mean"), scannf = FALSE, scale = FALSE)
s.class(x.pca$li, pop(gendata_Country), col = hybridpal()(11))

```
We see that some samples that have been assigned as wolves with SNP analyses, still point to dogs.

#3.2.2. Exploration with only Dinaric and Balkan samples

```{r}
gendata_Country <- gendata_PCA[pop(gendata_PCA) %in% c ("Dog", "W-Slovenia", "W-Croatia", "W-BIH", "W-Serbia", "W-Kosovo", "W-Montenegro", "W-N.Macedonia", "W-Bulgaria")]

x.pca <- dudi.pca(tab(gendata_Country, NA.method = "mean"), scannf = FALSE, scale = FALSE)
s.class(x.pca$li, pop(gendata_Country), col = hybridpal()(10))

```
```{r}
gendata_Country <- gendata_PCA[pop(gendata_PCA) %in% c ("Dog", "W-Slovenia", "W-Croatia", "W-N.Macedonia", "W-Bulgaria")]

x.pca <- dudi.pca(tab(gendata_Country, NA.method = "mean"), scannf = FALSE, scale = FALSE)
s.class(x.pca$li, pop(gendata_Country), col = hybridpal()(5))

```
A gradual cline between North Dinaric samples and South Balkan samples can bee seen.

#3.2.3) Exploration of dogs and samples from BIH, Srbia, Kosovo and Montenegro  - countries that showed mixed genes in preliminary STRUCTURE analyses from North Dinaric subpopulation (Slovenia, Croatia) and South Balkan subpopulation (North Macedonia, Bulgaria).

```{r}
gendata_Country <- gendata_PCA[pop(gendata_PCA) %in% c ("Dog", "W-BIH", "W-Serbia", "W-Kosovo", "W-Montenegro")]

x.pca <- dudi.pca(tab(gendata_Country, NA.method = "mean"), scannf = FALSE, scale = FALSE)
s.class(x.pca$li, pop(gendata_Country), col = hybridpal()(5))
```
Samples from BIH, Serbia, Kosovo and Montenegro cluster together and separately from dogs. Because of this we combined them together as "W-Middle_Balkan" samples in STRUCTURE analyses. 



#3.2.4) Dogs and other wolf samples

Italian samples
```{r}
gendata_Country <- gendata_PCA[pop(gendata_PCA) %in% c ("Dog",  "W-Italy")]

x.pca <- dudi.pca(tab(gendata_Country, NA.method = "mean"), scannf = FALSE, scale = FALSE)
s.class(x.pca$li, pop(gendata_Country), col = hybridpal()(2))
```
German samples
```{r}
gendata_Country <- gendata_PCA[pop(gendata_PCA) %in% c ("Dog",  "W-Germany")]

x.pca <- dudi.pca(tab(gendata_Country, NA.method = "mean"), scannf = FALSE, scale = FALSE)
s.class(x.pca$li, pop(gendata_Country), col = hybridpal()(2))
```
Romanian samples
```{r}
gendata_Country <- gendata_PCA[pop(gendata_PCA) %in% c ("Dog",  "W-Romania")]

x.pca <- dudi.pca(tab(gendata_Country, NA.method = "mean"), scannf = FALSE, scale = FALSE)
s.class(x.pca$li, pop(gendata_Country), col = hybridpal()(2))
```


4. Looking at Expected vs. Observed heterozygosity

4.1 Preparing data
#Combining BIH, Serbia, Kosovo and Montenegro together for analyses
```{r}
gendata_combined <- gendata_PCA

# Convert pop to character to allow new value assignment
pops <- as.character(pop(gendata_combined))

# Replace BIH, Serbia, Kosovo and Montenegro with "W-Middle Balkan"
pops[pops %in% c("W-BIH", "W-Serbia", "W-Kosovo", "W-Montenegro")] <- "W-Middle Balkan"

# Reassign and re-factor
pop(gendata_combined) <- as.factor(pops)

# Check population composition
table(pop(gendata_combined))
levels(pop(gendata_combined))
```


#Removing the hybrids
```{r}
hybrid_classes <- c("F1", "BC1w", "BC2w")
gendata_combined_H_removed <- gendata_combined[!pop(gendata_combined) %in% hybrid_classes, ]
gendata_combined_H_removed@other <- gendata_combined@other[indNames(gendata_combined), ]

# Check
table(pop(gendata_combined_H_removed))
levels(pop(gendata_combined_H_removed))
```




4.2 Calculate Ho and He statistics for every country for each locus and plot the data
```{r}
#Calculate Ho and He statistics for every population for each locus
set.seed(123)
Locus_statistic <- popSummLoc(gendata_combined_H_removed)

```


```{r}
#Ploting for all countries separate without correction

# Create a new column for outlier flag
Locus_statistic <- Locus_statistic %>%
  mutate(HWE_diff = abs(Ho - He),
         is_outlier = HWE_diff > 0.1) #Leniant outliers - highlighting clear outliers

# Plot, labeling only outliers
ggplot(Locus_statistic, aes(Ho, He, col = Locus)) +
  geom_point() +
  geom_text(data = filter(Locus_statistic, is_outlier),
            aes(label = Locus),
            vjust = -0.5, hjust = 0.5,
            size = 3, show.legend = FALSE) +
  facet_grid(Population ~ .) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Observed heterozygosity (Ho)", y = "Expected heterozygosity (He)")

```
4.3 Correcting the results 
We applied the Benjamini-Hochberg correction to control the False Discovery Rate (FDR). This adjustment prevents excessive false positives arising from multiple testing while maintaining higher statistical power. It ensures that the identified genetic variants represent true biological signals rather than statistical noise.

```{r}
Locus_statistic_corrected <- Locus_statistic %>%
  mutate(
    BH = p.adjust(Pr.exact, method = "BH") # correction with Benjamini-Hochberg (BH) procedure
  )

```


```{r}
#Corrected with BH
Corrected_BH <- Locus_statistic_corrected %>%
  mutate(is_significant = BH < 0.05)

HWB_BH_plot <- ggplot(Corrected_BH, aes(Ho, He, col = Locus)) +
  geom_point() +
  geom_text(data = filter(Corrected_BH, is_significant),
            aes(label = Locus),
            vjust = -0.5, hjust = 0.5,
            size = 3, show.legend = FALSE) +
  facet_grid(Population ~ .) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "black") +
  scale_x_continuous(limits = c(0, 1)) +
  scale_y_continuous(limits = c(0, 1)) +
  labs(x = "Observed heterozygosity (Ho)", y = "Expected heterozygosity (He)")

HWB_BH_plot

ggsave("Results_of_data_exploration/HWB_BH_plot_20MS.png", plot = HWB_BH_plot, width = 10, height = 10, dpi = 300)
```

5. Finding dog and wolf private alleles

5.1 Define populations and create datasets

```{r}
# =====================================================
# Define true wolf populations
# =====================================================

true_wolves <- c(
  "W-Slovenia",
  "W-Croatia",
  "W-Middle Balkan",
  "W-N.Macedonia",
  "W-Bulgaria",
  "W-Italy",
  "W-Germany",
  "W-Romania"
)


# =====================================================
# Create Dog + true wolf dataset
# =====================================================

g_dog_wolves <- gendata_combined[
  pop(gendata_combined) %in% c("Dog", true_wolves)
]

# Check populations included
table(pop(g_dog_wolves))


# =====================================================
# Create hybrid dataset
# =====================================================

hybrids <- c(
  "F1",
  "BC1w",
  "BC2w"
)

g_hybrids <- gendata_combined[
  pop(gendata_combined) %in% hybrids
]

# Check hybrid classes
table(pop(g_hybrids))
```

5.2 Identify dog- and wolf-private alleles

Private alleles are defined as alleles present in at least one individual of one group and absent from the other group.

```{r}
# =====================================================
# Allele presence table: Dog + true wolves
# =====================================================

all_tab <- tab(
  g_dog_wolves,
  freq = TRUE,
  NA.method = "zero"
)

dog_inds <- pop(g_dog_wolves) == "Dog"

wolf_inds <- pop(g_dog_wolves) %in% true_wolves


# Number of individuals carrying each allele
# (> 0 means the allele is present in that individual)

dog_presence <- colSums(
  all_tab[dog_inds, ] > 0
)

wolf_presence <- colSums(
  all_tab[wolf_inds, ] > 0
)


# =====================================================
# Dog-private alleles
# =====================================================

dog_private_alleles <- names(
  dog_presence[
    dog_presence > 0 &
    wolf_presence == 0
  ]
)


# =====================================================
# True wolf-private alleles
# =====================================================

wolf_private_alleles <- names(
  wolf_presence[
    wolf_presence > 0 &
    dog_presence == 0
  ]
)


# Results
dog_private_alleles
length(dog_private_alleles)

wolf_private_alleles
length(wolf_private_alleles)
```

5.3 Count individuals carrying each private allele

```{r}
# =====================================================
# Dog allele counts
# =====================================================

g_dogs <- gendata_combined[
  pop(gendata_combined) == "Dog"
]

dog_tab <- tab(
  g_dogs,
  freq = TRUE,
  NA.method = "zero"
)

dog_counts <- colSums(
  dog_tab[, dog_private_alleles, drop = FALSE] > 0
)


# =====================================================
# Wolf allele counts
# =====================================================

g_wolves <- gendata_combined[
  pop(gendata_combined) %in% true_wolves
]

wolf_tab <- tab(
  g_wolves,
  freq = TRUE,
  NA.method = "zero"
)

wolf_counts <- colSums(
  wolf_tab[, wolf_private_alleles, drop = FALSE] > 0
)


# =====================================================
# Identify private alleles present in only ONE
# individual
# =====================================================

dog_private_once <- names(
  dog_counts[
    dog_counts == 1
  ]
)

wolf_private_once <- names(
  wolf_counts[
    wolf_counts == 1
  ]
)


# Check numbers
length(dog_private_once)
length(wolf_private_once)
```

5.4 Identify dog-private alleles carried by hybrids

```{r}
# =====================================================
# Allele table for hybrids
# =====================================================

hyb_tab <- tab(
  g_hybrids,
  freq = TRUE,
  NA.method = "zero"
)


# =====================================================
# Number of hybrids carrying each dog-private allele
# =====================================================

hyb_counts <- colSums(
  hyb_tab[, dog_private_alleles, drop = FALSE] > 0
)
```

Exact dog-private alleles detected in each hybrid

```{r}
# =====================================================
# Exact dog-private alleles detected in each hybrid
# =====================================================

hybrid_private_exact <- lapply(
  seq_len(nInd(g_hybrids)),
  function(i) {
    
    ind_gt <- hyb_tab[i, ]
    
    found <- names(ind_gt)[
      names(ind_gt) %in% dog_private_alleles &
      ind_gt > 0
    ]
    
    tibble(
      SampleID = indNames(g_hybrids)[i],
      Pop = as.character(pop(g_hybrids)[i]),
      Dog_private_alleles = if (
        length(found) == 0
      ) {
        NA_character_
      } else {
        paste(found, collapse = ";")
      }
    )
  }
) %>%
  bind_rows()


hybrid_private_exact
```



5.5 Create per-allele summary information

```{r}
# =====================================================
# DOG PRIVATE ALLELE TABLE
# =====================================================

dog_private_table <- tibble(
  Allele = dog_private_alleles,
  N_Dogs = dog_counts[dog_private_alleles],
  N_Hybrids = hyb_counts[dog_private_alleles],
  Present_once_in_dog = dog_private_alleles %in% dog_private_once
)


# =====================================================
# WOLF PRIVATE ALLELE TABLE
# =====================================================

wolf_private_table <- tibble(
  Allele = wolf_private_alleles,
  N_Wolves = wolf_counts[wolf_private_alleles],
  Present_once_in_wolf = wolf_private_alleles %in% wolf_private_once
)


# Inspect
dog_private_table
wolf_private_table
```

5.6 Create locus-level summaries

```{r}
# =====================================================
# DOG PRIVATE ALLELES — LOCUS SUMMARY
# =====================================================

dog_locus <- dog_private_table %>%
  separate(
    Allele,
    into = c("Locus", "AlleleSize"),
    sep = "\\.",
    remove = TRUE
  ) %>%
  group_by(Locus) %>%
  summarise(
    Dog_private_n = n(),
    Dog_private_once_n = sum(Present_once_in_dog),
    Max_N_Dogs = max(N_Dogs),
    .groups = "drop"
  )


# =====================================================
# WOLF PRIVATE ALLELES — LOCUS SUMMARY
# =====================================================

wolf_locus <- wolf_private_table %>%
  separate(
    Allele,
    into = c("Locus", "AlleleSize"),
    sep = "\\.",
    remove = TRUE
  ) %>%
  group_by(Locus) %>%
  summarise(
    Wolf_private_n = n(),
    Wolf_private_once_n = sum(Present_once_in_wolf),
    Max_Wolves_private = max(N_Wolves),
    .groups = "drop"
  )


# =====================================================
# DOG-PRIVATE ALLELES DETECTED IN HYBRIDS
# =====================================================

hybrid_locus <- dog_private_table %>%
  separate(
    Allele,
    into = c("Locus", "AlleleSize"),
    sep = "\\.",
    remove = TRUE
  ) %>%
  group_by(Locus) %>%
  summarise(
    N_detected_in_hybrids = sum(N_Hybrids > 0),
    Max_N_Hybrids = max(N_Hybrids),
    .groups = "drop"
  )
```


5.7 Create the final locus-level table

This includes **all 20 loci**, including loci with zero private alleles.

```{r}
# =====================================================
# All loci
# =====================================================

all_loci <- locNames(gendata_combined)


# =====================================================
# FINAL SUMMARY TABLE
# =====================================================

private_alleles_locus_summary <- tibble(
  Locus = all_loci
) %>%
  
  left_join(
    dog_locus,
    by = "Locus"
  ) %>%
  
  left_join(
    wolf_locus,
    by = "Locus"
  ) %>%
  
  left_join(
    hybrid_locus,
    by = "Locus"
  ) %>%
  
  mutate(
    across(
      -Locus,
      ~ replace_na(., 0)
    )
  ) %>%
  
  select(
    Locus,
    Dog_private_n,
    Dog_private_once_n,
    Max_N_Dogs,
    Wolf_private_n,
    Wolf_private_once_n,
    Max_Wolves_private,
    N_detected_in_hybrids,
    Max_N_Hybrids
  )


private_alleles_locus_summary
```


5.8 Create overall summary statistics

```{r}
# =====================================================
# Overall numbers
# =====================================================

n_dog_private_alleles <- length(
  dog_private_alleles
)

n_wolf_private_alleles <- length(
  wolf_private_alleles
)


# Unique dog-private alleles detected in hybrids
alleles_in_hybrids <- hybrid_private_exact$Dog_private_alleles %>%
  na.omit() %>%
  strsplit(";") %>%
  unlist()

n_dog_private_in_hybrids <- length(
  unique(alleles_in_hybrids)
)


# Proportion of dog-private alleles detected in hybrids
prop_private_in_hybrids <-
  n_dog_private_in_hybrids /
  n_dog_private_alleles


# Summary
summary_private <- tibble(
  Total_dog_private_alleles = n_dog_private_alleles,
  Total_wolf_private_alleles = n_wolf_private_alleles,
  Detected_dog_private_in_hybrids = n_dog_private_in_hybrids,
  Proportion_dog_private_in_hybrids = prop_private_in_hybrids
)

summary_private
```

5.9 Export results to Excel

```{r}
# =====================================================
# Export all results to one Excel workbook
# =====================================================

wb <- createWorkbook()


# Locus-level summary
addWorksheet(
  wb,
  "Locus_summary"
)

writeData(
  wb,
  "Locus_summary",
  private_alleles_locus_summary
)


# Hybrid-specific results
addWorksheet(
  wb,
  "Hybrid_private_alleles"
)

writeData(
  wb,
  "Hybrid_private_alleles",
  hybrid_private_exact
)


# Dog private alleles
addWorksheet(
  wb,
  "Dog_private_alleles"
)

writeData(
  wb,
  "Dog_private_alleles",
  dog_private_table
)


# Wolf private alleles
addWorksheet(
  wb,
  "Wolf_private_alleles"
)

writeData(
  wb,
  "Wolf_private_alleles",
  wolf_private_table
)


# Overall summary
addWorksheet(
  wb,
  "Overall_summary"
)

writeData(
  wb,
  "Overall_summary",
  summary_private
)


# Formatting
for (sheet in names(wb)) {
  
  setColWidths(
    wb,
    sheet,
    cols = 1:ncol(
      read.xlsx(
        wb,
        sheet = sheet
      )
    ),
    widths = "auto"
  )
  
  freezePane(
    wb,
    sheet,
    firstRow = TRUE
  )
}


# Save
saveWorkbook(
  wb,
  file = "Results_of_data_exploration/Private_alleles_Dog_Wolf_Hybrids.xlsx",
  overwrite = TRUE
)
```



