1. Importing libraries, functions, samples and creating genind data
```{r}
#importing libraries
library(readxl)
library(tidyverse)
library(adegenet)

```

```{r}
# Functions for exporting genind datasets to STRUCTURE format
# Creates:
# 1) STRUCTURE genotype file (.txt)
# 2) Population ID legend (_POP.txt)
# 3) Sample ID–population mapping file (_PopMap.txt)

writeStructure <- function (genotypes,pop,path){
#STRUCTURE export
#genotypes = genotypes genind object
#pop = ordered population factor used for STRUCTURE population IDs
#path = output file path
  gen <- genind2df(genotypes,sep=" ")
  gen[is.na(gen)] <- "-9 -9"
  output <- cbind(id = row.names(gen),pop = as.numeric(pop),gen[2:ncol(gen)])
  cat(locNames(genotypes),"\n",file = path)
  write.table(output,path,sep = " ",row.names = FALSE,col.names = FALSE,append = TRUE,quote = FALSE)
  outLegend <- data.frame(id = 1:length(levels(pop)),Pop = levels(pop))
  print(outLegend)
  return(outLegend)
}

#
export_structure_dataset <- function(gen_sorted, out_prefix) {
#Exporting all three STRUCTURE-related files
#gen_sorted = sorted genind object with populations ordered according to STRUCTURE output requirements
#out_prefix = full path and file prefix for exported STRUCTURE files
  
    
  # Ensure output directory exists
  dir.create(dirname(out_prefix), recursive = TRUE, showWarnings = FALSE)
  
  # 1) STRUCTURE genotype file 
  strLegend <- writeStructure(
    gen_sorted,
    pop(gen_sorted),
    paste0(out_prefix, ".txt")
  )
  # 2) Population ID legend
  colnames(strLegend) <- c("PopID", "PopName")
  strLegend$PopName <- paste0(levels(pop(gen_sorted))[strLegend$PopID])

  write.table(
    strLegend,
    paste0(out_prefix, "_POP.txt"),
    sep = " ",
    row.names = FALSE,
    col.names = FALSE,
    quote = FALSE
  )
  
  # 3) Sample ID–population mapping
  sample_ids <- indNames(gen_sorted)
  pop_ids <- pop(gen_sorted)

  pop_df <- data.frame(
    SampleID = sample_ids,
    PopID = as.integer(pop_ids)
  )

  final_df <- merge(pop_df, strLegend, by = "PopID")

  final_df <- final_df[
    order(match(final_df$SampleID, sample_ids)),
    c("SampleID", "PopName")
  ]

  write.table(
    final_df,
    paste0(out_prefix, "_PopMap.txt"),
    sep = "\t",
    row.names = FALSE,
    col.names = FALSE,
    quote = FALSE
  )
}
```

```{r}
#Importing cleaned data
raw_data_all = read_xlsx("Data/Sample_data.xlsx", sheet = "Samples_data") 
```

```{r}
#Removing "_" from alleles because STRUCTURE can not read the data with them.

locus_cols <- which(names(raw_data_all) == "Cl147"):which(names(raw_data_all) == "Lup23")

raw_data_all[locus_cols] <- lapply(raw_data_all[locus_cols], function(x) {
  gsub("_", "", as.character(x))
})
```

```{r}
#Creating genind object for STRUCTURE data preparation 
gendata_cleaned <-  df2genind(raw_data_all[c("Cl147",	"Cl233",	"Cl264",	"Cl274",	"Cl285",	"Cl291",	"Cl308",	"Cl330",	"Cl344","Cl366",	"Cl375",	"Cl408",	"Cl507",	"Cl527",	"Cl380","Lup02",	"Lup13",	"Lup20",	"Lup21",	"Lup23")], sep = " ", NA.char = "-9 -9")
gendata_cleaned@other <-  raw_data_all[,1:10]
indNames(gendata_cleaned) <- raw_data_all$SampleID
pop(gendata_cleaned) <-  raw_data_all$Classified_STRUCTURE
unique(pop(gendata_cleaned))
```

2. Preparing data for STRUCTURE analyses

```{r}
#Filtering out only North Dinaric wolf samples from Slovenia and Croatia, hybrid samples and dog samples
gendata_NDinaric <- gendata_cleaned[pop(gendata_cleaned) %in% c ("Dog", "F1", "BC1w", "BC2w", "W-Slovenia", "W-Croatia")]

```

2.1 Slovenia–Croatia dataset without reference populations
1a) Using only empirical data confirmed with SNP analyses

```{r}
#Filtering and keeping only samples that were analysed with SNP panel
keep_snp <- gendata_NDinaric@other$"SNP_pannel" != "No"

gendata_NDinaric_SNP <- gendata_NDinaric[keep_snp, ]
```

```{r}
# Exporting data for STRUCTURE and further analyses

  # ------------------------------------------------------------
  # POPULATION ORDER 
  # ------------------------------------------------------------

desired_order <- c("Dog", "F1", "BC1w", "BC2w", "W-Slovenia", "W-Croatia")

  # ------------------------------------------------------------
  # REORDERING
  # ------------------------------------------------------------

message("Processing empirical data (no simulations)")

# Make sure ordering matches desired_order
present <- sort(unique(as.character(pop(gendata_NDinaric_SNP))))
unknown <- setdiff(present, desired_order)
if (length(unknown) > 0) {
  stop("Pop labels not in desired_order: ", paste(unknown, collapse = ", "))
}

# Reorder
pop(gendata_NDinaric_SNP) <- factor(pop(gendata_NDinaric_SNP), levels = desired_order)

pops <- pop(gendata_NDinaric_SNP)
sort_index <- order(factor(pops, levels = desired_order))
gen_sorted <- gendata_NDinaric_SNP[sort_index, ]
pop(gen_sorted) <- factor(pop(gen_sorted), levels = desired_order)


  # ------------------------------------------------------------
  # EXPORT
  # ------------------------------------------------------------

out_dir <- "Data_for_Structure/1a_NDinaric_empirical"

out_prefix <- file.path(out_dir, "1a_NDinaric_empirical")

export_structure_dataset(gen_sorted, out_prefix)
```
1b) Using empirical data with additional dogs not confirmed by SNP analyses

```{r}

  # ------------------------------------------------------------
  # POPULATION ORDER 
  # ------------------------------------------------------------

desired_order <- c("Dog", "F1",  "BC1w", "BC2w", "W-Slovenia", "W-Croatia")

  # ------------------------------------------------------------
  # REORDERING
  # ------------------------------------------------------------

message("Processing empirical data (no simulations)")


# Make sure ordering matches desired_order
present <- sort(unique(as.character(pop(gendata_NDinaric))))
unknown <- setdiff(present, desired_order)
if (length(unknown) > 0) {
  stop("Pop labels not in desired_order: ", paste(unknown, collapse = ", "))
}

# Reorder
pop(gendata_NDinaric) <- factor(pop(gendata_NDinaric), levels = desired_order)

pops <- pop(gendata_NDinaric)
sort_index <- order(factor(pops, levels = desired_order))
gen_sorted <- gendata_NDinaric[sort_index, ]
pop(gen_sorted) <- factor(pop(gen_sorted), levels = desired_order)


  # ------------------------------------------------------------
  # EXPORT
  # ------------------------------------------------------------

out_dir <- "Data_for_Structure/1b_NDinaric_empirical_all_dogs"

out_prefix <- file.path(out_dir, "1b_NDinaric_empirical_all_dogs")

export_structure_dataset(gen_sorted, out_prefix)
```
1c) Simulations of dog samples added to balance the numbers

```{r}
# ============================================================
# LOOP OVER SEEDS
# ============================================================

seeds <- 1:10

for (seed in seeds) {

  message("======================================")
  message("Processing SEED ", seed)
  message("======================================")

  set.seed(seed)

  # ------------------------------------------------------------
  # FUNCTION: expand population using ALL × ALL hybridization
  # ------------------------------------------------------------

  expand_population <- function(pop_name, n, new_pop_name) {

    inds <- gendata_NDinaric[which(pop(gendata_NDinaric) == pop_name)]

    if (length(inds) == 0) {
      stop(paste("Population", pop_name, "not found"))
    }

    hybrid <- hybridize(
      inds,
      inds,
      n   = n,
      pop = new_pop_name
    )

    return(hybrid)
  }

  # ------------------------------------------------------------
  # DEFINE POPULATIONS TO SIMULATE
  # ------------------------------------------------------------

  pop_list <- list(
    list(name = "Dog",         n = 250, new_pop = "Sim-dog")
  )

  # ------------------------------------------------------------
  # RUN SIMULATIONS
  # ------------------------------------------------------------

  expanded_populations <- list()

  for (pop in pop_list) {
    message("Processing population: ", pop$name)

    expanded_populations[[pop$new_pop]] <- expand_population(
      pop_name     = pop$name,
      n            = pop$n,
      new_pop_name = pop$new_pop
    )
  }

    if (length(expanded_populations) == 1) {
  simulations_dogs <- expanded_populations[[1]]
} else {
  simulations_dogs <- do.call(repool, expanded_populations)
}


  # ------------------------------------------------------------
  # COMBINE ORIGINAL DATA + SIMULATED DATA
  # ------------------------------------------------------------

  combined_wolves <- repool(
    gendata_NDinaric,
    simulations_dogs
  )

  # ------------------------------------------------------------
  # POPULATION ORDER 
  # ------------------------------------------------------------

  desired_order <- c(
    "Dog", "Sim-dog",
    "F1", "BC1w", "BC2w",
    "W-Slovenia", "W-Croatia"
  )

# Make sure ordering matches desired_order
present <- sort(unique(as.character(pop(combined_wolves))))
unknown <- setdiff(present, desired_order)
if (length(unknown) > 0) {
  stop("Pop labels not in desired_order: ", paste(unknown, collapse = ", "))
}

# Reorder  
  pop(combined_wolves) <- factor(pop(combined_wolves), levels = desired_order)

  pops <- pop(combined_wolves)

  sort_index <- order(factor(pops, levels = desired_order))
  gen_sorted <- combined_wolves[sort_index, ]
  pop(gen_sorted) <- factor(pop(gen_sorted), levels = desired_order)

  # ------------------------------------------------------------
  # EXPORT FOR STRUCTURE
  # ------------------------------------------------------------

  out_dir <- paste0(
    "Data_for_Structure/1c_NDinaric_empirical_Sim_dog/",
    "1c_NDinaric_empirical_Sim_dog_", seed
  )  
  
  out_prefix <- file.path(
  out_dir,
  paste0("1c_NDinaric_empirical_Sim_dog_", seed)
)

export_structure_dataset(gen_sorted, out_prefix)
}
```

2.2 Population openness comparison without reference populations – separate analyses for open population (Croatia) vs. closed population (Italy)

2ai)   Croatia - reduced sample size (100 per parental group) and simulated hybrids

```{r}
# ============================================================
# LOOP OVER SEEDS 
# ============================================================

seeds <- 1:10

for (seed in seeds) {

  message("======================================")
  message("Processing SEED ", seed)
  message("======================================")

  set.seed(seed)

  # ------------------------------------------------------------
  # FUNCTION: make hybrids  
  # ------------------------------------------------------------

  make_hybrids <- function(Dogs, wolves, tag) {

    F1   <- hybridize(Dogs, wolves, n = 100, pop = paste0("Sim-F1_", tag))
    BC1d <- hybridize(Dogs, F1,     n = 100, pop = paste0("Sim-BC1d_", tag))
    BC2d <- hybridize(Dogs, BC1d,   n = 100, pop = paste0("Sim-BC2d_", tag))
    BC1w  <- hybridize(wolves, F1,   n = 100, pop = paste0("Sim-BC1w_", tag))
    BC2w <- hybridize(wolves, BC1w,  n = 100, pop = paste0("Sim-BC2w_", tag))

    return(list(BC2d = BC2d, BC1d = BC1d, F1 = F1, BC1w = BC1w, BC2w = BC2w))
  }

  # ------------------------------------------------------------
  # SELECT CROATIA WOLVES 
  # ------------------------------------------------------------

  CROwolves_all <- gendata_cleaned[pop = "W-Croatia"]
  CROwolves_sel <- CROwolves_all[sample(1:nInd(CROwolves_all), 100)]

  # ------------------------------------------------------------
  # FILTER ORIGINAL DATA 
  # ------------------------------------------------------------

  gendata_filtered <- repool(
    gendata_cleaned[pop = c("Dog", "F1", "BC1w", "BC2w")],
    CROwolves_sel
  )

  Dogs <- gendata_cleaned[pop = "Dog"]

  # ------------------------------------------------------------
  # CREATE HYBRIDS 
  # ------------------------------------------------------------

  hybrids_CRO <- make_hybrids(Dogs, CROwolves_all, "CRO")

  combined_simulations_only_hybrids <- repool(
    hybrids_CRO$BC2d,
    hybrids_CRO$BC1d,
    hybrids_CRO$F1,
    hybrids_CRO$BC1w,
    hybrids_CRO$BC2w
  )

  # ------------------------------------------------------------
  # COMBINE ORIGINAL + SIMULATIONS 
  # ------------------------------------------------------------

  combined_all <- repool(
    gendata_filtered,
    combined_simulations_only_hybrids
  )

  # ------------------------------------------------------------
  # POPULATION ORDER 
  # ------------------------------------------------------------

  desired_order <- c(
    "Dog",
    "Sim-BC2d_CRO", "Sim-BC1d_CRO",
    "F1", "Sim-F1_CRO",
    "BC1w", "Sim-BC1w_CRO",
    "BC2w", "Sim-BC2w_CRO",
    "W-Croatia"
  )

  # Make sure ordering matches desired_order
  present <- sort(unique(as.character(pop(combined_all))))
  unknown <- setdiff(present, desired_order)

  if (length(unknown) > 0) {
    stop("Pop labels not in desired_order: ", paste(unknown, collapse = ", "))
  }
  
# Reorder
  pop(combined_all) <- factor(pop(combined_all), levels = desired_order)

  pops <- pop(combined_all)

  sort_index <- order(factor(pops, levels = desired_order))
  gen_sorted <- combined_all[sort_index, ]
  pop(gen_sorted) <- factor(pop(gen_sorted), levels = desired_order)



  # ------------------------------------------------------------
  # EXPORT FOR STRUCTURE 
  # ------------------------------------------------------------
  out_dir  <- paste0("Data_for_Structure/2ai_Croatia_low_numbers/", "2ai_Croatia_low_numbers_", seed)
    
  out_prefix <- file.path(out_dir, paste0("2ai_Croatia_low_numbers_", seed))

export_structure_dataset(gen_sorted, out_prefix)
}
```
2aii)  Italy - reduced sample size (100 per parental group) and simulated hybrids

```{r}
# ============================================================
# LOOP OVER SEEDS 
# ============================================================

seeds <- 1:10

for (seed in seeds) {

  message("======================================")
  message("Processing SEED ", seed)
  message("======================================")

  set.seed(seed)

  # ------------------------------------------------------------
  # KEEP ONLY EMPIRICAL SAMPLES 
  # ------------------------------------------------------------

  gendata_filtered <- gendata_cleaned[pop = c("Dog", "W-Italy")]

  # ------------------------------------------------------------
  # FUNCTION: expand population 
  # ------------------------------------------------------------

  expand_population <- function(pop_name, n, new_pop_name) {

    inds <- gendata_filtered[which(pop(gendata_filtered) == pop_name)]

    if (length(inds) == 0) {
      stop(paste("Population", pop_name, "not found"))
    }

    hybrid <- hybridize(
      inds,
      inds,
      n   = n,
      pop = new_pop_name
    )

    return(hybrid)
  }

  # ------------------------------------------------------------
  # DEFINE POPULATIONS TO EXPAND 
  # ------------------------------------------------------------

  pop_list <- list(
    list(name = "W-Italy",  n = 72, new_pop = "Sim-W-Italy")
  )

  # ------------------------------------------------------------
  # RUN EXPANSIONS 
  # ------------------------------------------------------------

  expanded_populations <- list()

  for (pop_item in pop_list) {

    message("Expanding population: ", pop_item$name)

    expanded_populations[[pop_item$new_pop]] <- expand_population(
      pop_name     = pop_item$name,
      n            = pop_item$n,
      new_pop_name = pop_item$new_pop
    )
  }

  if (length(expanded_populations) == 1) {
  combined_simulations_no_hybrids <- expanded_populations[[1]]
} else {
  combined_simulations_no_hybrids <- do.call(repool, expanded_populations)
}

  # ------------------------------------------------------------
  # FUNCTION: make hybrids 
  # ------------------------------------------------------------

  make_hybrids <- function(Dogs, wolves, tag) {

    F1   <- hybridize(Dogs, wolves, n = 100, pop = paste0("Sim-F1_", tag))
    BC1d <- hybridize(Dogs, F1,     n = 100, pop = paste0("Sim-BC1d_", tag))
    BC2d <- hybridize(Dogs, BC1d,   n = 100, pop = paste0("Sim-BC2d_", tag))
    BC1w  <- hybridize(wolves, F1,   n = 100, pop = paste0("Sim-BC1w_", tag))
    BC2w <- hybridize(wolves, BC1w,  n = 100, pop = paste0("Sim-BC2w_", tag))

    return(list(BC2d = BC2d, BC1d = BC1d, F1 = F1, BC1w = BC1w, BC2w = BC2w))
  }

  # ------------------------------------------------------------
  # CREATE ITALIAN HYBRIDS 
  # ------------------------------------------------------------

  Dogs      <- gendata_filtered[pop = "Dog"]
  ITAwolves <- gendata_filtered[pop = "W-Italy"]

  hybrids_ITA <- make_hybrids(Dogs, ITAwolves, "ITA")

  combined_simulations_only_hybrids <- repool(
    hybrids_ITA$BC2d,
    hybrids_ITA$BC1d,
    hybrids_ITA$F1,
    hybrids_ITA$BC1w,
    hybrids_ITA$BC2w
  )

  # ------------------------------------------------------------
  # COMBINE ORIGINAL + SIMULATIONS 
  # ------------------------------------------------------------

  combined_all <- repool(
    gendata_filtered,
    combined_simulations_only_hybrids,
    combined_simulations_no_hybrids
  )

  # ------------------------------------------------------------
  # POPULATION ORDER 
  # ------------------------------------------------------------

  desired_order <- c(
  "Dog",
  "Sim-BC2d_ITA", "Sim-BC1d_ITA", 
  "Sim-F1_ITA", "Sim-BC1w_ITA", "Sim-BC2w_ITA",
  "W-Italy", "Sim-W-Italy"
  )

# Make sure ordering matches desired_order
  present <- sort(unique(as.character(pop(combined_all))))
  unknown <- setdiff(present, desired_order)

  if (length(unknown) > 0) {
    stop("Pop labels not in desired_order: ", paste(unknown, collapse = ", "))
  }

# Reorder
  pop(combined_all) <- factor(pop(combined_all), levels = desired_order)

  pops <- pop(combined_all)

  sort_index <- order(factor(pops, levels = desired_order))
  gen_sorted <- combined_all[sort_index, ]
  pop(gen_sorted) <- factor(pop(gen_sorted), levels = desired_order)


  # ------------------------------------------------------------
  # EXPORT FOR STRUCTURE 
  # ------------------------------------------------------------
  out_dir  <- paste0("Data_for_Structure/2aii_Italy_low_numbers/","2aii_Italy_low_numbers_", seed)
  
  out_prefix <- file.path(
    out_dir,
    paste0("2aii_Italy_low_numbers_", seed)
  )

export_structure_dataset(gen_sorted, out_prefix)
}
```

2bi)   Croatia - increased sample size (500 per parental group) and simulated hybrids
```{r}
# ============================================================
# LOOP OVER SEEDS 
# ============================================================

seeds <- 1:10

for (seed in seeds) {

  message("======================================")
  message("Processing SEED ", seed)
  message("======================================")

  set.seed(seed)

  # ------------------------------------------------------------
  # KEEP ONLY EMPIRICAL SAMPLES 
  # ------------------------------------------------------------

  gendata_filtered <- gendata_cleaned[pop = c("Dog", "W-Croatia", "F1", "BC1w", "BC2w")]

  # ------------------------------------------------------------
  # FUNCTION: expand population 
  # ------------------------------------------------------------

  expand_population <- function(pop_name, n, new_pop_name) {

    inds <- gendata_filtered[which(pop(gendata_filtered) == pop_name)]

    if (length(inds) == 0) {
      stop(paste("Population", pop_name, "not found"))
    }

    hybrid <- hybridize(
      inds,
      inds,
      n   = n,
      pop = new_pop_name
    )

    return(hybrid)
  }

  # ------------------------------------------------------------
  # DEFINE POPULATIONS TO EXPAND 
  # ------------------------------------------------------------

  pop_list <- list(
    list(name = "Dog",      n = 400, new_pop = "Sim-dog"),
    list(name = "W-Croatia",  n = 260, new_pop = "Sim-W-Croatia")
  )

  # ------------------------------------------------------------
  # RUN EXPANSIONS 
  # ------------------------------------------------------------

  expanded_populations <- list()

  for (pop_item in pop_list) {

    message("Expanding population: ", pop_item$name)

    expanded_populations[[pop_item$new_pop]] <- expand_population(
      pop_name     = pop_item$name,
      n            = pop_item$n,
      new_pop_name = pop_item$new_pop
    )
  }

  combined_simulations_no_hybrids <- do.call(repool, expanded_populations)

  # ------------------------------------------------------------
  # FUNCTION: make hybrids 
  # ------------------------------------------------------------

  make_hybrids <- function(Dogs, wolves, tag) {

    F1   <- hybridize(Dogs, wolves, n = 100, pop = paste0("Sim-F1_", tag))
    BC1d <- hybridize(Dogs, F1,     n = 100, pop = paste0("Sim-BC1d_", tag))
    BC2d <- hybridize(Dogs, BC1d,   n = 100, pop = paste0("Sim-BC2d_", tag))
    BC1w  <- hybridize(wolves, F1,   n = 100, pop = paste0("Sim-BC1w_", tag))
    BC2w <- hybridize(wolves, BC1w,  n = 100, pop = paste0("Sim-BC2w_", tag))

    return(list(BC2d = BC2d, BC1d = BC1d, F1 = F1, BC1w = BC1w, BC2w = BC2w))
  }

  # ------------------------------------------------------------
  # CREATE CROATIA HYBRIDS 
  # ------------------------------------------------------------

  Dogs      <- gendata_filtered[pop = "Dog"]
  CROwolves <- gendata_filtered[pop = "W-Croatia"]

  hybrids_CRO <- make_hybrids(Dogs, CROwolves, "CRO")

  combined_simulations_only_hybrids <- repool(
    hybrids_CRO$BC2d,
    hybrids_CRO$BC1d,
    hybrids_CRO$F1,
    hybrids_CRO$BC1w,
    hybrids_CRO$BC2w
  )

  # ------------------------------------------------------------
  # COMBINE ORIGINAL + SIMULATIONS 
  # ------------------------------------------------------------

  combined_all <- repool(
    gendata_filtered,
    combined_simulations_only_hybrids,
    combined_simulations_no_hybrids
  )

  # ------------------------------------------------------------
  # POPULATION ORDER 
  # ------------------------------------------------------------

  desired_order <- c(
    "Dog", "Sim-dog",
    "Sim-BC2d_CRO", "Sim-BC1d_CRO",
    "F1", "Sim-F1_CRO",
    "BC1w", "Sim-BC1w_CRO",
    "BC2w", "Sim-BC2w_CRO",
    "W-Croatia", "Sim-W-Croatia"
  )

# Make sure ordering matches desired_order
  present <- sort(unique(as.character(pop(combined_all))))
  unknown <- setdiff(present, desired_order)

  if (length(unknown) > 0) {
    stop("Pop labels not in desired_order: ", paste(unknown, collapse = ", "))
  }

# Reorder
  pop(combined_all) <- factor(pop(combined_all), levels = desired_order)

  pops <- pop(combined_all)

  sort_index <- order(factor(pops, levels = desired_order))
  gen_sorted <- combined_all[sort_index, ]
  pop(gen_sorted) <- factor(pop(gen_sorted), levels = desired_order)



  # ------------------------------------------------------------
  # EXPORT FOR STRUCTURE 
  # ------------------------------------------------------------
  
  out_dir  <- paste0("Data_for_Structure/2bi_Croatia_high_numbers/","2bi_Croatia_high_numbers_", seed)
  out_prefix <- file.path(
    out_dir,
    paste0("2bi_Croatia_high_numbers_", seed)
  )

export_structure_dataset(gen_sorted, out_prefix)
}
```

2bii)  Italy - increased sample size (500 per parental group) and simulated hybrids

```{r}
# ============================================================
# LOOP OVER SEEDS 
# ============================================================

seeds <- 1:10

for (seed in seeds) {

  message("======================================")
  message("Processing SEED ", seed)
  message("======================================")

  set.seed(seed)

  # ------------------------------------------------------------
  # KEEP ONLY EMPIRICAL SAMPLES 
  # ------------------------------------------------------------

  gendata_filtered <- gendata_cleaned[pop = c("Dog", "W-Italy")]

  # ------------------------------------------------------------
  # FUNCTION: expand population 
  # ------------------------------------------------------------

  expand_population <- function(pop_name, n, new_pop_name) {

    inds <- gendata_filtered[which(pop(gendata_filtered) == pop_name)]

    if (length(inds) == 0) {
      stop(paste("Population", pop_name, "not found"))
    }

    hybrid <- hybridize(
      inds,
      inds,
      n   = n,
      pop = new_pop_name
    )

    return(hybrid)
  }

  # ------------------------------------------------------------
  # DEFINE POPULATIONS TO EXPAND 
  # ------------------------------------------------------------

  pop_list <- list(
    list(name = "Dog",      n = 400, new_pop = "Sim-dog"),
    list(name = "W-Italy",  n = 472, new_pop = "Sim-W-Italy")
  )

  # ------------------------------------------------------------
  # RUN EXPANSIONS 
  # ------------------------------------------------------------

  expanded_populations <- list()

  for (pop_item in pop_list) {

    message("Expanding population: ", pop_item$name)

    expanded_populations[[pop_item$new_pop]] <- expand_population(
      pop_name     = pop_item$name,
      n            = pop_item$n,
      new_pop_name = pop_item$new_pop
    )
  }

  combined_simulations_no_hybrids <- do.call(repool, expanded_populations)

  # ------------------------------------------------------------
  # FUNCTION: make hybrids 
  # ------------------------------------------------------------

  make_hybrids <- function(Dogs, wolves, tag) {

    F1   <- hybridize(Dogs, wolves, n = 100, pop = paste0("Sim-F1_", tag))
    BC1d <- hybridize(Dogs, F1,     n = 100, pop = paste0("Sim-BC1d_", tag))
    BC2d <- hybridize(Dogs, BC1d,   n = 100, pop = paste0("Sim-BC2d_", tag))
    BC1w  <- hybridize(wolves, F1,   n = 100, pop = paste0("Sim-BC1w_", tag))
    BC2w <- hybridize(wolves, BC1w,  n = 100, pop = paste0("Sim-BC2w_", tag))

    return(list(BC2d = BC2d, BC1d = BC1d, F1 = F1, BC1w = BC1w, BC2w = BC2w))
  }

  # ------------------------------------------------------------
  # CREATE ITALIAN HYBRIDS 
  # ------------------------------------------------------------

  Dogs      <- gendata_filtered[pop = "Dog"]
  ITAwolves <- gendata_filtered[pop = "W-Italy"]

  hybrids_ITA <- make_hybrids(Dogs, ITAwolves, "ITA")

  combined_simulations_only_hybrids <- repool(
    hybrids_ITA$BC2d,
    hybrids_ITA$BC1d,
    hybrids_ITA$F1,
    hybrids_ITA$BC1w,
    hybrids_ITA$BC2w
  )

  # ------------------------------------------------------------
  # COMBINE ORIGINAL + SIMULATIONS 
  # ------------------------------------------------------------

  combined_all <- repool(
    gendata_filtered,
    combined_simulations_only_hybrids,
    combined_simulations_no_hybrids
  )

  # ------------------------------------------------------------
  # POPULATION ORDER 
  # ------------------------------------------------------------

desired_order <- c(
  "Dog","Sim-dog",
  "Sim-BC2d_ITA", "Sim-BC1d_ITA", 
  "Sim-F1_ITA", "Sim-BC1w_ITA", "Sim-BC2w_ITA",
  "W-Italy", "Sim-W-Italy"
)

# Make sure ordering matches desired_order
  present <- sort(unique(as.character(pop(combined_all))))
  unknown <- setdiff(present, desired_order)

  if (length(unknown) > 0) {
    stop("Pop labels not in desired_order: ", paste(unknown, collapse = ", "))
  }
  
# Reorder
  pop(combined_all) <- factor(pop(combined_all), levels = desired_order)

  pops <- pop(combined_all)

  sort_index <- order(factor(pops, levels = desired_order))
  gen_sorted <- combined_all[sort_index, ]
  pop(gen_sorted) <- factor(pop(gen_sorted), levels = desired_order)


  # ------------------------------------------------------------
  # EXPORT FOR STRUCTURE 
  # ------------------------------------------------------------

  out_dir  <- paste0("Data_for_Structure/2bii_Italy_high_numbers/","2bii_Italy_high_numbers_", seed)
  
  out_prefix <- file.path(
    out_dir,
    paste0("2bii_Italy_high_numbers_", seed)
  )

export_structure_dataset(gen_sorted, out_prefix)
}

```


2.3 All empirical samples with reference populations

3a) Empirical individuals only 

```{r}

  # ------------------------------------------------------------
  # POPULATION ORDER 
  # ------------------------------------------------------------

desired_order <- c("Dog", "Sim-dog","F1", "BC1w","BC2w", "W-Slovenia", "W-Croatia", "W-Middle_Balkan", "W-N.Macedonia", "W-Bulgaria", "W-Italy", "W-Germany", "W-Romania")

  # ------------------------------------------------------------
  # REORDERING
  # ------------------------------------------------------------

message("Processing empirical data (no simulations)")

# Make sure ordering matches desired_order
present <- sort(unique(as.character(pop(gendata_cleaned))))
unknown <- setdiff(present, desired_order)
if (length(unknown) > 0) {
  stop("Pop labels not in desired_order: ", paste(unknown, collapse = ", "))
}

# Reorder
pop(gendata_cleaned) <- factor(pop(gendata_cleaned), levels = desired_order)

pops <- pop(gendata_cleaned)
sort_index <- order(factor(pops, levels = desired_order))
gen_sorted <- gendata_cleaned[sort_index, ]
pop(gen_sorted) <- factor(pop(gen_sorted), levels = desired_order)


  # ------------------------------------------------------------
  # EXPORT
  # ------------------------------------------------------------

out_dir <- "Data_for_Structure/3a_Empirical_samples_no_sim"

out_prefix <- file.path(out_dir, "3a_Empirical_samples_no_sim")

export_structure_dataset(gen_sorted, out_prefix)

```
3b) Empirical individuals and simulated reference samples to balance groups sizes 

```{r}
# ============================================================
# LOOP OVER SEEDS
# ============================================================

seeds <- 1:10

for (seed in seeds) {

  message("======================================")
  message("Processing SEED ", seed)
  message("======================================")

  set.seed(seed)

  # ------------------------------------------------------------
  # FUNCTION: expand population using ALL × ALL hybridization
  # ------------------------------------------------------------

  expand_population <- function(pop_name, n, new_pop_name) {

    inds <- gendata_cleaned[which(pop(gendata_cleaned) == pop_name)]

    if (length(inds) == 0) {
      stop(paste("Population", pop_name, "not found"))
    }

    hybrid <- hybridize(
      inds,
      inds,
      n   = n,
      pop = new_pop_name
    )

    return(hybrid)
  }

  # ------------------------------------------------------------
  # DEFINE POPULATIONS TO SIMULATE
  # ------------------------------------------------------------

  pop_list <- list(
    list(name = "Dog",         n = 250, new_pop = "Sim-dog"),
    list(name = "W-Germany",     n = 300, new_pop = "Sim-W-Germany"),
    list(name = "W-Italy",       n = 300, new_pop = "Sim-W-Italy"),
    list(name = "W-Romania",     n = 300, new_pop = "Sim-W-Romania"),
    list(name = "W-N.Macedonia", n = 125, new_pop = "Sim-W-N.Macedonia"),
    list(name = "W-Bulgaria",    n = 125, new_pop = "Sim-W-Bulgaria")
  )

  # ------------------------------------------------------------
  # RUN SIMULATIONS
  # ------------------------------------------------------------

  expanded_populations <- list()

  for (pop in pop_list) {
    message("Processing population: ", pop$name)

    expanded_populations[[pop$new_pop]] <- expand_population(
      pop_name     = pop$name,
      n            = pop$n,
      new_pop_name = pop$new_pop
    )
  }

  combined_simulations_no_hybrids <- do.call(repool, expanded_populations)

  # ------------------------------------------------------------
  # COMBINE ORIGINAL DATA + SIMULATED DATA
  # ------------------------------------------------------------

  combined_wolves <- repool(
    gendata_cleaned,
    combined_simulations_no_hybrids
  )

  # ------------------------------------------------------------
  # POPULATION ORDER 
  # ------------------------------------------------------------

  desired_order <- c(
    "Dog", "Sim-dog",
    "F1", "BC1w", "BC2w",
    "W-Slovenia", "W-Croatia", "W-Middle_Balkan",
    "W-N.Macedonia", "Sim-W-N.Macedonia",
    "W-Bulgaria", "Sim-W-Bulgaria",
    "W-Italy", "Sim-W-Italy",
    "W-Germany", "Sim-W-Germany",
    "W-Romania", "Sim-W-Romania"
  )

# Make sure ordering matches desired_order
present <- sort(unique(as.character(pop(combined_wolves))))
unknown <- setdiff(present, desired_order)
if (length(unknown) > 0) {
  stop("Pop labels not in desired_order: ", paste(unknown, collapse = ", "))
}

# Reorder
  pop(combined_wolves) <- factor(pop(combined_wolves), levels = desired_order)

  pops <- pop(combined_wolves)

  sort_index <- order(factor(pops, levels = desired_order))
  gen_sorted <- combined_wolves[sort_index, ]
  pop(gen_sorted) <- factor(pop(gen_sorted), levels = desired_order)


  # ------------------------------------------------------------
  # EXPORT FOR STRUCTURE
  # ------------------------------------------------------------
  out_dir <- paste0(
    "Data_for_Structure/3b_Empirical_Simulation_data_no_hybrids/",
    "3b_Empirical_Simulation_data_no_hybrids_", seed
  )
  
  out_prefix <- file.path(out_dir, paste0("3b_Empirical_Simulation_data_no_hybrids_", seed))

export_structure_dataset(gen_sorted, out_prefix)
}
```
2.4. All empirical samples with reference populations and simulated hybrids of different categories

4a) Empirical individuals, additional simulated reference samples to achieve balanced sample sizes and simulated hybrids from five clusters 

```{r}
# ============================================================
# LOOP OVER SEEDS
# ============================================================

seeds <- 1:10

for (seed in seeds) {

  message("======================================")
  message("Processing SEED ", seed)
  message("======================================")

  set.seed(seed)

  # ------------------------------------------------------------
  # FUNCTION: expand population using ALL × ALL hybridization
  # ------------------------------------------------------------

  expand_population <- function(pop_name, n, new_pop_name) {

    inds <- gendata_cleaned[which(pop(gendata_cleaned) == pop_name)]

    if (length(inds) == 0) {
      stop(paste("Population", pop_name, "not found"))
    }

    hybrid <- hybridize(
      inds,
      inds,
      n   = n,
      pop = new_pop_name
    )

    return(hybrid)
  }

  # ------------------------------------------------------------
  # FUNCTION: make hybrids 
  # ------------------------------------------------------------

  make_hybrids <- function(Dogs, wolves, tag) {

    F1   <- hybridize(Dogs, wolves, n = 100, pop = paste0("Sim-F1_", tag))
    BC1w  <- hybridize(wolves, F1,   n = 100, pop = paste0("Sim-BC1w_", tag))
    BC2w <- hybridize(wolves, BC1w,  n = 100, pop = paste0("Sim-BC2w_", tag))

    return(list(F1 = F1, BC1w = BC1w, BC2w = BC2w))
  }

  # ------------------------------------------------------------
  # DEFINE POPULATIONS TO SIMULATE 
  # ------------------------------------------------------------

  pop_list <- list(
    list(name = "Dog",         n = 200, new_pop = "Sim-dog"),
    list(name = "W-Germany",     n = 300, new_pop = "Sim-W-Germany"),
    list(name = "W-Italy",       n = 300, new_pop = "Sim-W-Italy"),
    list(name = "W-Romania",     n = 300, new_pop = "Sim-W-Romania"),
    list(name = "W-N.Macedonia", n = 125, new_pop = "Sim-W-N.Macedonia"),
    list(name = "W-Bulgaria",    n = 125, new_pop = "Sim-W-Bulgaria")
  )

  # ------------------------------------------------------------
  # RUN SIMULATIONS 
  # ------------------------------------------------------------

  expanded_populations <- list()

  for (pop_item in pop_list) {
    message("Processing population: ", pop_item$name)

    expanded_populations[[pop_item$new_pop]] <- expand_population(
      pop_name     = pop_item$name,
      n            = pop_item$n,
      new_pop_name = pop_item$new_pop
    )
  }

  combined_simulations_no_hybrids <- do.call(repool, expanded_populations)

  # ------------------------------------------------------------
  # CREATE HYBRIDS
  # ------------------------------------------------------------

  Dogs       <- gendata_cleaned[which(pop(gendata_cleaned) == "Dog")]
  CROwolves  <- gendata_cleaned[which(pop(gendata_cleaned) == "W-Croatia")]
  BULwolves  <- gendata_cleaned[which(pop(gendata_cleaned) == "W-Bulgaria")]
  ITAwolves  <- gendata_cleaned[which(pop(gendata_cleaned) == "W-Italy")]
  GERwolves  <- gendata_cleaned[which(pop(gendata_cleaned) == "W-Germany")]
  ROMwolves  <- gendata_cleaned[which(pop(gendata_cleaned) == "W-Romania")]

  hybrids_CRO <- make_hybrids(Dogs, CROwolves, "CRO")
  hybrids_BUL <- make_hybrids(Dogs, BULwolves, "BUL")
  hybrids_ITA <- make_hybrids(Dogs, ITAwolves, "ITA")
  hybrids_GER <- make_hybrids(Dogs, GERwolves, "GER")
  hybrids_ROM <- make_hybrids(Dogs, ROMwolves, "ROM")

  combined_simulations_only_hybrids <- repool(
    hybrids_CRO$F1, hybrids_CRO$BC1w, hybrids_CRO$BC2w,
    hybrids_BUL$F1, hybrids_BUL$BC1w, hybrids_BUL$BC2w,
    hybrids_ITA$F1, hybrids_ITA$BC1w, hybrids_ITA$BC2w,
    hybrids_GER$F1, hybrids_GER$BC1w, hybrids_GER$BC2w,
    hybrids_ROM$F1, hybrids_ROM$BC1w, hybrids_ROM$BC2w
  )

  # ------------------------------------------------------------
  # COMBINE ORIGINAL + ALL SIMULATIONS
  # ------------------------------------------------------------

  combined_all <- repool(
    gendata_cleaned,
    combined_simulations_no_hybrids,
    combined_simulations_only_hybrids
  )

  # ------------------------------------------------------------
  # POPULATION ORDER 
  # ------------------------------------------------------------

  desired_order <- c(
    "Dog","Sim-dog",
    "F1","Sim-F1_CRO","Sim-F1_BUL","Sim-F1_ITA","Sim-F1_GER","Sim-F1_ROM",
    "BC1w","Sim-BC1w_CRO","Sim-BC1w_BUL","Sim-BC1w_ITA","Sim-BC1w_GER","Sim-BC1w_ROM",
    "BC2w","Sim-BC2w_CRO","Sim-BC2w_BUL","Sim-BC2w_ITA","Sim-BC2w_GER","Sim-BC2w_ROM",
    "W-Slovenia","W-Croatia","W-Middle_Balkan",
    "W-N.Macedonia","Sim-W-N.Macedonia",
    "W-Bulgaria","Sim-W-Bulgaria",
    "W-Italy","Sim-W-Italy",
    "W-Germany","Sim-W-Germany",
    "W-Romania","Sim-W-Romania"
  )

  # Make sure ordering matches desired_order
present <- sort(unique(as.character(pop(combined_all))))
unknown <- setdiff(present, desired_order)
if (length(unknown) > 0) {
  stop("Pop labels not in desired_order: ", paste(unknown, collapse = ", "))
}

# Reorder
  pop(combined_all) <- factor(pop(combined_all), levels = desired_order)

  pops <- pop(combined_all)

  sort_index <- order(factor(pops, levels = desired_order))
  gen_sorted <- combined_all[sort_index, ]
  pop(gen_sorted) <- factor(pop(gen_sorted), levels = desired_order)


  # ------------------------------------------------------------
  # EXPORT FOR STRUCTURE 
  # ------------------------------------------------------------
  out_dir <- paste0(
    "Data_for_Structure/4a_Simulation_data_wolves_hybrids/",
    "4a_Simulation_data_wolves_hybrids_", seed
  )
  
  out_prefix <- file.path(
    out_dir,
    paste0("4a_Simulation_data_wolves_hybrids_", seed)
  )

export_structure_dataset(gen_sorted, out_prefix)
}
```

