Code for running ParallesStructure on linux

# Template: analyses with simulation replicates
```{r}
# ============================================================
# Run STRUCTURE analyses using ParallelStructure
# ============================================================

# ---------- Load library and set path ----------
# Install and load ParallelStructure package
#install.packages("ParallelStructure", repos="http://R-Forge.R-project.org") #only run the first time
library(ParallelStructure)

# Path to the STRUCTURE installation.
# CHANGE this to the location of the STRUCTURE executable.
structure_path <- "/path/to/Structure/"

# ---------- Function to run one job ----------
run_job <- function(jobnum) {
  # Path to the ParallelStructure joblist file. 
  # CHANGE the path and filename pattern to match your files.
joblist <- paste0(
  "/path/to/analysis_folder/dataset_",
  jobnum,
  "/joblist_",
  jobnum,
  ".txt"
)

infile <- paste0(
  "/path/to/analysis_folder/dataset_",
  jobnum,
  "/dataset_",
  jobnum,
  ".txt"
)

outpath <- paste0(
  "/path/to/analysis_folder/dataset_",
  jobnum,
  "/"
)

  # Check that required files exist
if (!file.exists(joblist)) {
  stop("Joblist file not found: ", joblist)
}

if (!file.exists(infile)) {
  stop("STRUCTURE input file not found: ", infile)
}  

  # ---------- Run ParallelStructure ----------
  
  parallel_structure(
    joblist = joblist,
    n_cpu = 20,              # Number of CPUs to use. Change according to available resources.
    structure_path = structure_path,
    infile = infile,
    outpath = outpath,
    
    numinds = 745,          # Number of individuals in the dataset. CHANGE to match your dataset.
    numloci = 20,            # Number of loci/markers. CHANGE to match your dataset.
    
    plot_output = 1,         # Generate STRUCTURE plot output.
    label = 1,               # Input file contains individual labels.
    popdata = 1,             # Input file contains population information.
    popflag = 0,             # Population flag is not used.
    phenotypes = 0,          # No phenotype data.
    markernames = 1,         # Input file contains marker/locus names.
    mapdist = 0,             # No map distances are used.
    onerowperind = 1,        # Each individual is represented by one row.
    phaseinfo = 0,           # No phase information.
    extracol = 0,            # No extra columns.
    
    missing = -9,            # Missing genotype value used in the input file.
    ploidy = 2,              # Ploidy of the organisms.
    
    usepopinfo = 0,          # Disable USEPOPINFO model; population labels are not used as priors.
    
    revert_convert = 1,      # Convert STRUCTURE output back to the original format.
    printqhat = 1,           # Print individual Q estimates.
    
    locdata = 0,             # No location information.
    recessivealleles = 0,    # No recessive alleles.
    phased = 0,               # Data are not phased.
    noadmix = 0,             # Use the admixture model.
    linkage = 0,              # Markers are treated as unlinked.
    locprior = 0,             # Do not use the LOCPRIOR model.
    
    inferalpha = 1           # Estimate alpha from the data.
  )
}

# ---------- Run jobs sequentially ----------

# Specify the jobs to run.
# For example:
#   1:10       = run jobs 1 through 10
#   c(1,3,7)   = run only jobs 1, 3 and 7
#
# CHANGE this to the jobs you want to analyse.

for (i in 1:10) {
  cat("Running job", i, "...\n")
  run_job(i)
  cat("Job", i, "finished.\n\n")
}


```

#EXAMPLES
#Example 1: Simulated datasets with simulation replicates (1-10) specific folders (analyses 1c, 2ai, 2aii, 2bi, 2bii, 3b, 4a).
Example for analyses 1c, where folders are arranged as (each simulation replicate folder contains a STRUCTURE input file and a joblist defining the STRUCTURE runs):

 1c_NDinaric_empirical_Sim_dog/
  │
  ├── 1c_NDinaric_empirical_Sim_dog_1/
  │   ├── joblist_1c_1.txt
  │   ├── 1c_NDinaric_empirical_Sim_dog_1.txt
  │   ├── ...
  │
  ├── 1c_NDinaric_empirical_Sim_dog_2/
  │   ├── joblist_1c_2.txt
  │   ├── 1c_NDinaric_empirical_Sim_dog_2.txt
  │   ├── ...
  │...


```{r}

# ---------- Load library and set path ----------
library(ParallelStructure)

structure_path <- "/path/to/Structure/"

# ---------- Function to run one job ----------
run_job <- function(jobnum) {
  joblist <- paste0("/ParallelStructure_Results/1c_NDinaric_empirical_Sim_dog/1c_NDinaric_empirical_Sim_dog_", jobnum, "/joblist_1c_", jobnum, ".txt")
  infile  <- paste0("/ParallelStructure_Results/1c_NDinaric_empirical_Sim_dog/1c_NDinaric_empirical_Sim_dog_", jobnum, "/1c_NDinaric_empirical_Sim_dog_", jobnum, ".txt")
  outpath <- paste0("/ParallelStructure_Results/1c_NDinaric_empirical_Sim_dog/1c_NDinaric_empirical_Sim_dog_", jobnum, "/")
  
    # Check that required files exist
if (!file.exists(joblist)) {
  stop("Joblist file not found: ", joblist)
}

if (!file.exists(infile)) {
  stop("STRUCTURE input file not found: ", infile)
} 
  
  parallel_structure(
    joblist = joblist,
    n_cpu = 5, # Number of CPUs/processes used by ParallelStructure; here set to match the number of STRUCTURE runs in the joblist. CHANGE if needed.
    structure_path = structure_path,
    infile = infile,
    outpath = outpath,
    numinds = 745,  #CHANGE to match your dataset.
    numloci = 20,  #CHANGE to match your dataset.
    plot_output = 1,
    label = 1,
    popdata = 1,
    popflag = 0,
    phenotypes = 0,
    markernames = 1,
    mapdist = 0,
    onerowperind = 1,
    phaseinfo = 0,
    extracol = 0,
    missing = -9,
    ploidy = 2,
    usepopinfo = 0,
    revert_convert = 1,
    printqhat = 1,
    locdata = 0,
    recessivealleles = 0,
    phased = 0,
    noadmix = 0,
    linkage = 0,
    locprior = 0,
    inferalpha = 1
  )
}

# ---------- Run jobs sequentially ----------
for (i in 1:10) {
  cat("Running job", i, "...\n")
  run_job(i)
  cat("Job", i, "finished.\n\n")
}


```

#Example 2: Single STRUCTURE run without simulation replicates (analyses 1a, 1b, 3a).
Example for analyses 3a.
```{r}

# ---------- Load library and set path ----------
library(ParallelStructure)

structure_path <- "/path/to/Structure/"

# ---------- One job run without function ----------
  joblist <- "/ParallelStructure_Results/3a_Empirical_samples_no_sim/joblist_3a.txt"
  infile  <- "/ParallelStructure_Results/3a_Empirical_samples_no_sim/3a_Empirical_samples_no_sim.txt"
  outpath <- "/ParallelStructure_Results/3a_Empirical_samples_no_sim/"
  
  parallel_structure(
    joblist = joblist,
    n_cpu = 20,  #CHANGE if needed
    structure_path = structure_path,
    infile = infile,
    outpath = outpath,
    numinds = 742,  #CHANGE to match your dataset.
    numloci = 20,  #CHANGE to match your dataset.
    plot_output = 1,
    label = 1,
    popdata = 1,
    popflag = 0,
    phenotypes = 0,
    markernames = 1,
    mapdist = 0,
    onerowperind = 1,
    phaseinfo = 0,
    extracol = 0,
    missing = -9,
    ploidy = 2,
    usepopinfo = 0,
    revert_convert = 1,
    printqhat = 1,
    locdata = 0,
    recessivealleles = 0,
    phased = 0,
    noadmix = 0,
    linkage = 0,
    locprior = 0,
    inferalpha = 1
  )

```
